Making Sequential Programs Deterministic

### 2019-04-06 design.tex

Suppose that $\<myStrings> : \ptype{OrderNonDet List}{Det String}$.
The type of \<first(myStrings)> should be \<NonDet String>.

@@

\noindent
would be invalid.
At the call \<first(myStrings)>, the most specific instantiation would be

@@

The rules prevent some invocations.  Suppose we have the following variable:

$\<listOfLists> : \ptype{OrderNonDet List}{\ptype{Det List}{Det String}}$

@@

\todo{TODO: Need to work on this section.}

In the annotated JDK, the type of get is:

\begin{Verbatim}
    @PolyDet("up") E get(@PolyDet List<E> this, @PolyDet int index);
\end{Verbatim}

@@

$\<get> : \forallt{\kappa} \kappa\ \<List>\angles{\kappa\ \<String>} \times \kappa\ \<Int> \rightarrow \kappa\ \<String>$
\todo{Used to be $\up{\kappa}$ for return type, but I don't think that is needed.}
\todo{Oh, maybe it's because we need to be able to instantiate in the tye
  parameter differently?  How do we enable that?}

(The actual typing permits the qualifier on the type argument to differ
from the type qualifier on the list, and it handles arbitrary basetypes
including collections that may themselves be order-nondeterministic.)

Instantiations are:

@@

\todo{We don't need the \<OrderNonDet> instantiation.  The other two
  instantiations are enough.
  We don't need the following instantiation:}
\begin{Verbatim}
   // instantiation: OrderNonDet List<Det String> x _ Int -> NonDet String 
\end{Verbatim}

If either the list or the index is (order) non-deterministic, then the result is
nondeterministic:

\begin{Verbatim}
NonDet Int ni;
   Det Int di;
NonDet List<NonDet String> nl;
OrderNonDet List<Det String> ol;
Det List<Det String> dl;

get(nl, ni) : NonDet String // instantiation: NonDet List<NonDet String> x NonDet Int -> NonDet String
get(nl, di) : NonDet String
get(ol, ni) : NonDet String
get(ol, di) : NonDet String
get(dl, ni) : NonDet String
get(dl, di) : Det String    // instantiation: NonDet List<Det String> x Det Int -> Det String
\end{Verbatim}

>>>>

Suppose that $\<myOndStrings> : \ptype{OrderNonDet List}{Det String}$.
The type of \<first(myOndStrings)> should be \<NonDet String>.

@@

\noindent
would be invalid.
At the call \<first(myOndStrings)>, the most specific instantiation would be

@@

The given type prevents some invocations.  Suppose we have the following variable:

$\<listOfLists> : \ptype{OrderNonDet List}{\lptype{Det List}{Det String}}$

@@

This section introduces the another polymorphism feature by contrasting the
types of the \<get> and \<set> operations on lists.

@@

$\<get> : \forallt{\kappa} \kappa\ \<List>\angles{\kappa\ \<String>} \times \kappa\ \<Int> \rightarrow \up{\kappa}\ \<String>$

\noindent
Its instantiations are:

@@

\noindent
It cannot be instantiated at $\kappa = \<OrderNonDet>$, but the other two
instantiations are enough to ensure that if either the list or the index is (order) non-deterministic, then the result is
nondeterministic.

% \begin{Verbatim}
% NonDet Int ni;
%    Det Int di;
% NonDet List<NonDet String> nl;
% OrderNonDet List<Det String> ol;
% Det List<Det String> dl;
% 
% get(nl, ni) : NonDet String // instantiation: NonDet List<NonDet String> x NonDet Int -> NonDet String
% get(nl, di) : NonDet String
% get(ol, ni) : NonDet String
% get(ol, di) : NonDet String
% get(dl, ni) : NonDet String
% get(dl, di) : Det String    // instantiation: NonDet List<Det String> x Det Int -> Det String
% \end{Verbatim}
